### 2024-09-24 paper/paper.qmd
Laid out structure for paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: Emily Su
thanks: "Code and data are available at: https://github.com/moonsdust/toronto-homicides-shootings."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}
## Methodology

## Data Source and Measurements

## Variables of Interest

# Results {#sec-results}
This is the results section

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Areas of improvement

Weaknesses and next steps should also be included.

## Next steps

\newpage

\appendix

# Appendix {-}

## Acknowledgments 
We would like the acknowledge @citetellingstorieswithdata for some R code that 
was used to produce the tables and graphs.

## Note on Reproducing 
To reproduce the results in the paper, first run the scripts found in the scripts folder of the GitHub repository corresponding to the paper starting with the script, 00-install_packages.R to install the necessary packages.

## Code styling 
Code written in the scripts was checked and styled with lintr [@citelintr] and styler [@citestyler].

## Additional Figures

## Additional Tables

\newpage

# References

### 2024-09-24 scripts/02-data_cleaning.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Cleans the unedited ward profile data, unedited police location
-# data, and unedited data on neighbourhood crime rates. 
+# data, and unedited data on neighbourhood crime rates.
 # Author: Emily Su
 # Date: 21 September 2024
 # Contact: em.su@mail.utoronto.ca
@@ -31,8 +31,8 @@ raw_neighbourhood_data <-
 cleaned_ward_data <- raw_ward_profiles_data |> clean_names()
 
 # Get row that contains the average household income
-cleaned_ward_data <- 
-  cleaned_ward_data |> 
+cleaned_ward_data <-
+  cleaned_ward_data |>
   filter(x2021_one_variable_city_of_toronto_profiles == 
            "Average total income of households in 2020 ($)")
 # Convert row to a column 

### 2024-09-24 paper/paper.qmd
Updated subtitle and title of paper

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Toronto Homicides and Shootings (To be Changed)"
+subtitle: "An analysis of police presence across Toronto neighbourhoods in relation to homicide and shooting cases (2018-2023)"
 author: Emily Su

### 2024-09-25 scripts/00-install_packages.R
Added geojsonsf package

@@ -17,3 +17,4 @@ install.packages("dplyr")
 install.packages("ggplot2") # To make graphs
 install.packages("lintr") # To check styling of code
 install.packages("styler") # To style code
+install.packages("geojsonsf") # To convert from GeoJSON to sf object

### 2024-09-25 scripts/00-install_packages.R
Added sf package

@@ -17,4 +17,5 @@ install.packages("dplyr")
 install.packages("ggplot2") # To make graphs
 install.packages("lintr") # To check styling of code
 install.packages("styler") # To style code
-install.packages("geojsonsf") # To convert from GeoJSON to sf object
+install.packages("geojsonsf") 
+install.packages("sf")

### 2024-09-25 scripts/01-download_data.R
Updated download and cleaning scripts + datasets

@@ -25,23 +25,33 @@
 library(opendatatoronto)
 library(tidyverse)
 library(dplyr)
+library(sf)
 
 #### Download data ####
 # Code to download data was adapted from:
 # https://open.toronto.ca/dataset/neighbourhood-crime-rates/
 
+# # Function to download data
+# download_csv <- function(package_num) {
+#   files <- filter(list_package_resources(package_num),
+#                   tolower(format) %in% c("csv"))
+#   return(filter(files, row_number() == 1) %>% get_resource())
+# }
+
 # Function to download data
-download_csv <- function(package_num) {
+# Downloading geojson file due to the formatting of "geometry" column
+download_geojson <- function(package_num) {
   files <- filter(list_package_resources(package_num),
-                  tolower(format) %in% c("csv"))
+                  tolower(format) %in% c("geojson"))
   return(filter(files, row_number() == 1) %>% get_resource())
 }
 
 # Download "Neighbourhood Crime Rates" Dataset
-crime_rates <- download_csv("neighbourhood-crime-rates")
+# crime_rates <- download_csv("neighbourhood-crime-rates")
+crime_rates <- download_geojson("neighbourhood-crime-rates")
 
 # Download "Police Facility Locations" Dataset
-police_location <- download_csv("9aeefa17-27e8-4dd9-b74d-80f7f9eb85ac")
+police_location <- download_geojson("9aeefa17-27e8-4dd9-b74d-80f7f9eb85ac")
 
 # Download "Ward Profiles (25-Ward Model)"
 ward_profiles <-
@@ -49,11 +59,12 @@ ward_profiles <-
          row_number() == 1) |> get_resource()
 
 # Download "Neighbourhood"
-neighbourhood_boundaries <- download_csv("neighbourhoods")
+neighbourhood_boundaries <- download_geojson("neighbourhoods")
 
 #### Save data ####
-write_csv(crime_rates, "data/raw_data/raw_crime_rates.csv")
-write_csv(police_location, "data/raw_data/raw_police_location.csv")
+# write_csv(crime_rates, "data/raw_data/raw_crime_rates.csv")
+write_sf(crime_rates, "data/raw_data/raw_crime_rates.geojson")
+write_sf(police_location, "data/raw_data/raw_police_location.geojson")
 write_csv(as.data.frame(ward_profiles[1]),
           "data/raw_data/raw_ward_profiles.csv")
[…]

### 2024-09-25 scripts/02-data_cleaning.R

@@ -14,16 +14,24 @@
 library(tidyverse)
 library(janitor)
 library(dplyr)
+library(geojsonsf)
+library(sf)
 
 #### Clean data ####
 raw_ward_profiles_data <-
   read_csv("data/raw_data/raw_ward_profiles.csv", show_col_types = FALSE)
 raw_crime_rates_data <-
-  read_csv("data/raw_data/raw_crime_rates.csv", show_col_types = FALSE)
+  geojson_sf("data/raw_data/raw_crime_rates.geojson")
 raw_police_location_data <-
-  read_csv("data/raw_data/raw_police_location.csv", show_col_types = FALSE)
+  geojson_sf("data/raw_data/raw_police_location.geojson")
 raw_neighbourhood_data <-
-  read_csv("data/raw_data/raw_neighbourhood.csv", show_col_types = FALSE)
+  geojson_sf("data/raw_data/raw_neighbourhood.geojson")
+# raw_crime_rates_data <-
+#   read_csv("data/raw_data/raw_crime_rates.csv", show_col_types = FALSE)
+#raw_police_location_data <-
+#  read_csv("data/raw_data/raw_police_location.csv", show_col_types = FALSE)
+#raw_neighbourhood_data <-
+#  read_csv("data/raw_data/raw_neighbourhood.csv", show_col_types = FALSE)
 
 # Dataset 1 (Ward Profiles Dataset)
 # Expected columns: ward_num | ward_name | avg_income
@@ -100,6 +108,21 @@ cleaned_crime_data <-
          homicide_2020, homicide_2021, homicide_2022, homicide_2023, 
          shooting_2018, shooting_2019, shooting_2020, shooting_2021,
          shooting_2022, shooting_2023, geometry)
+
+# Convert all cases columns to numeric 
+cleaned_crime_data$homicide_2018<- as.numeric(cleaned_crime_data$homicide_2018)
+cleaned_crime_data$homicide_2019<- as.numeric(cleaned_crime_data$homicide_2019)
+cleaned_crime_data$homicide_2020 <- as.numeric(cleaned_crime_data$homicide_2020)
+cleaned_crime_data$homicide_2021 <- as.numeric(cleaned_crime_data$homicide_2021)
+cleaned_crime_data$homicide_2022 <- as.numeric(cleaned_crime_data$homicide_2022)
+cleaned_crime_data$homicide_2023 <- as.numeric(cleaned_crime_data$homicide_2023)
+
[…]

### 2024-09-25 scripts/01-download_data.R
Removed neighbourhood boundaries dataset

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from Open Data Toronto on the location
-# of police facilities, neighbourhood crime rates, the profiles of
-# Toronto's 25 wards, and neighbourhood boundaries.
+# of police facilities, neighbourhood crime rates, and the profiles of
+# Toronto's 25 wards.
 # Author: Emily Su
 # Date: 21 September 2024
 # Contact: em.su@mail.utoronto.ca
@@ -18,8 +18,6 @@
 # - https://open.toronto.ca/dataset/police-facility-locations/
 # Neighbourhood Crime Rates
 # - https://open.toronto.ca/dataset/neighbourhood-crime-rates/
-# Neighbourhoods
-# - https://open.toronto.ca/dataset/neighbourhoods/
 
 #### Workspace setup ####
 library(opendatatoronto)
@@ -58,13 +56,9 @@ ward_profiles <-
   filter(list_package_resources("6678e1a6-d25f-4dff-b2b7-aa8f042bc2eb"),
          row_number() == 1) |> get_resource()
 
-# Download "Neighbourhood"
-neighbourhood_boundaries <- download_geojson("neighbourhoods")
-
 #### Save data ####
 # write_csv(crime_rates, "data/raw_data/raw_crime_rates.csv")
 write_sf(crime_rates, "data/raw_data/raw_crime_rates.geojson")
 write_sf(police_location, "data/raw_data/raw_police_location.geojson")
 write_csv(as.data.frame(ward_profiles[1]),
           "data/raw_data/raw_ward_profiles.csv")
-write_sf(neighbourhood_boundaries, "data/raw_data/raw_neighbourhood.geojson")

### 2024-09-25 scripts/02-data_cleaning.R

@@ -24,14 +24,10 @@ raw_crime_rates_data <-
   geojson_sf("data/raw_data/raw_crime_rates.geojson")
 raw_police_location_data <-
   geojson_sf("data/raw_data/raw_police_location.geojson")
-raw_neighbourhood_data <-
-  geojson_sf("data/raw_data/raw_neighbourhood.geojson")
 # raw_crime_rates_data <-
 #   read_csv("data/raw_data/raw_crime_rates.csv", show_col_types = FALSE)
 #raw_police_location_data <-
 #  read_csv("data/raw_data/raw_police_location.csv", show_col_types = FALSE)
-#raw_neighbourhood_data <-
-#  read_csv("data/raw_data/raw_neighbourhood.csv", show_col_types = FALSE)
 
 # Dataset 1 (Ward Profiles Dataset)
 # Expected columns: ward_num | ward_name | avg_income
@@ -210,15 +206,6 @@ cleaned_police_location <- raw_police_location_data |> clean_names()
 cleaned_police_location <-
   cleaned_police_location |> select(facility, geometry)
 
-# Dataset 4
-# Expected Columns: area_short_code | area_name | geometry
-# Clean column names 
-cleaned_neighbourhood_data <- raw_neighbourhood_data |> clean_names()
-
-# Select expected columns 
-cleaned_neighbourhood_data <-
-  cleaned_neighbourhood_data |> select(area_short_code, area_name, geometry)
-
 #### Save data ####
 # Save crime data as a geojson
 write_sf(cleaned_crime_data, "data/analysis_data/cleaned_crime_data.geojson")
@@ -227,9 +214,6 @@ write_csv(cleaned_ward_data, "data/analysis_data/cleaned_ward_data.csv")
 # Save police location data as a geojson
 write_sf(cleaned_police_location,
           "data/analysis_data/cleaned_police_location.geojson")
-# Save neighbourhood data as a geojson
-write_sf(cleaned_neighbourhood_data,
-          "data/analysis_data/cleaned_neighbourhood.geojson")
 # # Save crime data as a CSV
 # write_csv(cleaned_crime_data, "data/analysis_data/cleaned_crime_data.csv")
 # # Save ward profile data as a CSV
@@ -237,6 +221,3 @@ write_sf(cleaned_neighbourhood_data,
 # # Save police location data as a CSV
 # write_csv(cleaned_police_location,
 #           "data/analysis_data/cleaned_police_location.csv")
[…]

### 2024-09-25 paper/paper.qmd
Finished graphs and updated paper

@@
 
-library(tidyverse)
+library("tidyverse") # Contains data-related packages
+library("knitr") # To make tables
+library("opendatatoronto")
+library("ggplot2") # To make graphs
+library("sf") # To make maps
+library("geojsonsf") # To read geojson files
+```
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Read in data
+# Police data 
+cleaned_police_location <- geojson_sf('../data/analysis_data/cleaned_police_location.geojson')
+# Crime data
+cleaned_crime_data <- geojson_sf('../data/analysis_data/cleaned_crime_data.geojson')
+# Ward data 
+cleaned_ward_data <- read_csv('../data/analysis_data/cleaned_ward_data.csv', show_col_types = FALSE)
 ```
@@
 # Results {#sec-results}
-This is the results section
+## Number of Homicide and Shooting Cases in Toronto from 2018 to 2023 in Toronto
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+# Add up total number of cases per neighbourhood
+crime_per_neighbourhood <- 
+  cleaned_crime_data |>
+  # Calculate the total sum of cases across all neighbourhoods 
+  mutate(
+    num_cases_total = sum(num_of_cases)
+  ) |>
+  select(neighbourhood, homicide_or_shooting, num_of_cases, num_cases_total) |>
+  # Groups data by type of crime, and neighbourhood
+  group_by(neighbourhood, homicide_or_shooting) |>
+  # Create new columns for number of cases for homicides and shootings per neighbourhood. Also calculate the proportion as well. 
+  mutate (
+    num_cases_total_per_neighbourhood_per_crime = sum(num_of_cases),
+    num_cases_total_per_neighbourhood_per_crime_proportion = round(num_cases_total_per_neighbourhood_per_crime / num_cases_total, 3)
+  ) |>
+  ungroup() |>
+  group_by(neighbourhood) |> 
+  mutate(
+    num_cases_total_per_neighbourhood = sum(num_of_cases),
+    num_cases_total_per_neighbourhood_proportion = round(num_cases_total_per_neighbourhood / num_cases_total, 3)
+  ) |>
+  # Obtains unique rows
+  unique() |>
+  # Arrange based on the total number of cases per neighbourhood
+  arrange(desc(num_cases_total_per_neighbourhood)) |>
+  tibble() |>
+  select(neighbourhood, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion)
+
+# Add up total number of cases each year across all neighbourhoods
+yearly_crime <- 
+  cleaned_crime_data |>
+  # Calculate the total sum of cases across all neighbourhoods 
+  mutate(
+    num_cases_total = sum(num_of_cases)
+  ) |>
+  # Group data by year, type of crime, number of total cases, and number of cases
+  select(year, homicide_or_shooting, num_cases_total, num_of_cases) |>
+  group_by(year, homicide_or_shooting) |>
+  unique() |>
+  # Create new columns for number of cases per year for homicides and shootings across ALL neighbourhood. Also calculate the proportion as well. 
+  mutate (
+    yearly_num_cases_total = sum(num_of_cases),
+    yearly_num_cases_total_proportion = round(yearly_num_cases_total / num_cases_total, 3)
+  ) |>
+  tibble() |>
+  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_proportion) |>
+  unique() |>
+  # Arrange based on the total number of cases per year 
+  arrange(desc(yearly_num_cases_total)) |>
+  ungroup()
+```
+
+```{r}
+#| label: tbl-totalcasestable
+#| fig-cap:  Number of Homicide and Shooting Cases in Toronto from 2018 to 2023
+#| echo: false
+#| warning: false
+#| message: false
+
+# Creates table
+yearly_crime |>
+  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_proportion) |>
+  # Arrange based on year
+  arrange(year) |>
+  unique() |>
+  rename(
+    "Type of crime" = homicide_or_shooting,
+    "Total number of cases yearly" = yearly_num_cases_total,
+    "Proportion of cases yearly" = yearly_num_cases_total_proportion,
+    "Year" = year
+  ) |>
+  kable()
+```
+
+```{r}
+#| label: fig-totalcaseshistogram
+#| fig-cap:  Number of Homicide and Shooting Cases in Toronto from 2018 to 2023
+#| echo: false
+#| warning: false
+#| message: false
+
+yearly_crime |> 
+  ggplot(mapping = aes(x = year, y = yearly_num_cases_total, fill = homicide_or_shooting)) +
+  geom_bar(stat="identity", position = "dodge2") +
+  labs(x = "Year", y = "Number of Cases", fill = "Type of Crime") +
+  theme_minimal() + 
+  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7)) + scale_fill_viridis_d()
+```
+
+```{r}
+#| label: tbl-totalcasesstats
+#| tbl-cap: Minimum, quartiles, median, and maximum of the number of homicides and shootings in Toronto from 2018 to 2023
+#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
+#| warning: false # Warnings are not included 
+#| message: false # Messages are not included
+
+# Creates table
+yearly_crime |>
+  select(yearly_num_cases_total) |>
+  rename(
+    "Total number of cases" = yearly_num_cases_total
+  ) |>
+  summary() |>
+  kable()
+```
+
+```{r}
+#| label: tbl-totalshootingsstats
+#| tbl-cap: Minimum, quartiles, median, and maximum of the number of shootings in Toronto from 2018 to 2023
+#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
+#| warning: false # Warnings are not included 
+#| message: false # Messages are not included
+
+# Creates table
+yearly_crime |>
+  filter(homicide_or_shooting == "shooting") |>
+  select(yearly_num_cases_total) |>
+  rename(
+    "Total number of cases" = yearly_num_cases_total
+  ) |>
+  summary() |>
+  kable()
+```
+
+```{r}
+#| label: tbl-totalhomicidesstats
+#| tbl-cap: Minimum, quartiles, median, and maximum of the number of homicides in Toronto from 2018 to 2023
+#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
+#| warning: false # Warnings are not included 
+#| message: false # Messages are not included
+
+# Creates table
+yearly_crime |>
+  filter(homicide_or_shooting == "homicide") |>
+  select(yearly_num_cases_total) |>
+  rename(
+    "Total number of cases" = yearly_num_cases_total
+  ) |>
+  summary() |>
+  kable()
+```
+
+## Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings
+
+```{r}
+#| label: tbl-highestcasesshootingstable
+#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Shootings from 2018 to 2023
+#| echo: false
+#| warning: false
+#| message: false
+
+# Creates table
+crime_per_neighbourhood |>
+  filter(homicide_or_shooting == "shooting") |>
+  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
+  select(neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
+  unique() |>
+  rename(
+    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
+    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
+    "Neighbourhood" = neighbourhood
+  ) |>
+  head(6) |>
+  kable()
+```
+
+```{r}
+#| label: tbl-highestcaseshomicidestable
+#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides from 2018 to 2023
+#| echo: false
+#| warning: false
+#| message: false
+
+# Creates table
+crime_per_neighbourhood |>
+  filter(homicide_or_shooting == "homicide") |>
+  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
+  select(neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
+  unique() |>
+  rename(
+    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
+    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
+    "Neighbourhood" = neighbourhood
+  ) |>
+  head(6) |>
+  kable()
+```
+
+```{r}
+#| label: tbl-highestcasestable
+#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings Combined from 2018 to 2023
+#| echo: false
+#| warning: false
+#| message: false
+
+# Creates table
+crime_per_neighbourhood |>
+  filter(homicide_or_shooting == "homicide") |>
+  arrange(desc(num_cases_total_per_neighbourhood)) |>
+  select(neighbourhood, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
+  unique() |>
+  rename(
+    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood,
+    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_proportion,
+    "Neighbourhood" = neighbourhood
+  ) |>
+  head(6) |>
+  kable()
+```
+
+```{r}
+#| label: fig-highestcaseshistogram
+#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings from 2018 to 2023
+#| echo: false
+#| warning: false
+#| message: false
+
+crime_per_neighbourhood |> 
+  select(neighbourhood, homicide_or_shooting, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime) |>
+  arrange(desc(num_cases_total_per_neighbourhood)) |>
+  select(-c(num_cases_total_per_neighbourhood)) |>
+  unique() |>
+  head(12) |>
+  ggplot(mapping = aes(x = neighbourhood, y = num_cases_total_per_neighbourhood_per_crime, fill = homicide_or_shooting)) +
+  geom_bar(stat="identity", position = "dodge2") +
+  labs(x = "Year", y = "Number of Cases", fill = "Type of Crime") +
+  theme_minimal() + 
+  # Reference https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2 to figure out how to rotate text along x axis.
+  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  + scale_fill_viridis_d()
+```
+
+## Police Facility Locations in relation to the Number of Homicides and Shootings from 2018 to 2023 in Toronto
+
+```{r}
+#| label: fig-homicides-map
+#| fig-cap: Location of police facilities in relation to Toronto neighbourhoods with the highest number of homicides from 2018 to 2023.
+#| echo: false
+#| warning: false
+#| message: false
+# Reference the following for code:
+# https://cengel.github.io/R-spatial/mapping.html
+ggplot() + 
+    # To help convert between GeoJSON to sf  https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html
+  # Fill in area regarding the number of cases 
+  geom_sf(data = cleaned_crime_data |> filter(homicide_or_shooting == "homicide"), aes(fill = num_of_cases, geometry = geometry)) +
+  # Changes the font of the title and removes the x and y axis 
+  # Reference to figured out how to add labels to maps https://tmieno2.github.io/R-as-GIS-for-Economists/create-maps.html
+  theme_minimal() + 
+  labs(
+    x = "Longitude",
+    y = "Latitude",
+    title = "Number of homicides across Toronto neighbourhoods (2018 to 2023)",
+    fill = "Number of cases"
+  ) +
+  geom_sf_text(
+    data = cleaned_crime_data |> filter(num_of_cases >= 14),
+    aes(label = hood_id),
+    check_overlap = TRUE,
+    size = 2, 
+    color = "white"
+    ) +
+    theme(plot.title = element_text(size = 9, face = 'bold'), 
+          plot.subtitle = element_text(size = 9),
+          rect = element_blank(), 
+          axis.text.x = element_blank(), 
+          axis.text.y = element_blank(), 
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(), 
+          axis.ticks = element_blank(), 
+          legend.title = element_text(size = 8),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank()
+          ) +
+  # Add in police locations
+  geom_point(data = cleaned_police_location,
+    aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
+  ) + labs(color = "Location")
+
+```
+
+```{r}
+#| label: fig-shootings-map
+#| fig-cap: Location of police facilities in relation to Toronto neighbourhoods with the highest number of shootings from 2018 to 2023.
+#| echo: false
+#| warning: false
+#| message: false
+# Reference the following for code:
+# https://cengel.github.io/R-spatial/mapping.html
+ggplot() + 
+    # To help convert between GeoJSON to sf  https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html
+  # Fill in area regarding the number of cases 
+  geom_sf(data = cleaned_crime_data |> filter(homicide_or_shooting == "shooting"), aes(fill = num_of_cases, geometry = geometry)) +
+  # Changes the font of the title and removes the x and y axis 
+  # Reference to figured out how to add labels to maps https://tmieno2.github.io/R-as-GIS-for-Economists/create-maps.html
+  theme_minimal() + 
+  labs(
+    x = "Longitude",
+    y = "Latitude",
+    title = "Number of shootings across Toronto neighbourhoods (2018 to 2023)",
+    fill = "Number of cases"
+  ) +
+  geom_sf_text(
+    data = cleaned_crime_data |> filter(num_of_cases >= 14),
+    aes(label = hood_id),
+    check_overlap = TRUE,
+    size = 2, 
+    color = "white"
+    ) +
+    theme(plot.title = element_text(size = 9, face = 'bold'), 
+          plot.subtitle = element_text(size = 9),
+          rect = element_blank(), 
+          axis.text.x = element_blank(), 
+          axis.text.y = element_blank(), 
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(), 
+          axis.ticks = element_blank(), 
+          legend.title = element_text(size = 8),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank()
+          ) +
+  # Add in police locations
+  geom_point(data = cleaned_police_location,
+    aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
+  ) + labs(color = "Location")
+
+```
 

### 2024-09-25 paper/paper.qmd
Updated graphs and tables in paper

@@
   ) |>
-  select(neighbourhood, homicide_or_shooting, num_of_cases, num_cases_total) |>
+  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total) |>
   # Groups data by type of crime, and neighbourhood
@@
   tibble() |>
-  select(neighbourhood, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion)
+  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion)
 
@@
 #| label: tbl-totalcasestable
-#| fig-cap:  Number of Homicide and Shooting Cases in Toronto from 2018 to 2023
+#| tbl-cap: Number of Homicide and Shooting Cases in Toronto from 2018 to 2023
 #| echo: false
@@
 #| label: tbl-highestcasesshootingstable
-#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Shootings from 2018 to 2023
+#| tbl-cap:  Six Toronto Neighbourhoods with the Highest Cases of Shootings from 2018 to 2023
 #| echo: false
@@
   arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
-  select(neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
+  select(neighbourhood, hood_id, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
   unique() |>
@@
     "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
+    "Neighbourhood ID" = hood_id,
     "Neighbourhood" = neighbourhood
@@
 #| label: tbl-highestcaseshomicidestable
-#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides from 2018 to 2023
+#| tbl-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides from 2018 to 2023
 #| echo: false
@@
     "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
+    "Neighbourhood ID" = hood_id, 
     "Neighbourhood" = neighbourhood
@@
 #| label: tbl-highestcasestable
-#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings Combined from 2018 to 2023
+#| tbl-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings Combined from 2018 to 2023
 #| echo: false
@@
   arrange(desc(num_cases_total_per_neighbourhood)) |>
-  select(neighbourhood, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
+  select(neighbourhood, hood_id, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
   unique() |>
@@
     "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_proportion,
+    "Neighbourhood ID" = hood_id,
     "Neighbourhood" = neighbourhood
@@
 crime_per_neighbourhood |> 
-  select(neighbourhood, homicide_or_shooting, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime) |>
+  select(hood_id, homicide_or_shooting, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime) |>
   arrange(desc(num_cases_total_per_neighbourhood)) |>
@@
   head(12) |>
-  ggplot(mapping = aes(x = neighbourhood, y = num_cases_total_per_neighbourhood_per_crime, fill = homicide_or_shooting)) +
+  ggplot(mapping = aes(x = as.character(hood_id), y = num_cases_total_per_neighbourhood_per_crime, fill = homicide_or_shooting)) +
   geom_bar(stat="identity", position = "dodge2") +
-  labs(x = "Year", y = "Number of Cases", fill = "Type of Crime") +
+  labs(x = "Neighbourhood ID", y = "Number of Cases", fill = "Type of Crime") +
   theme_minimal() + 
   # Reference https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2 to figure out how to rotate text along x axis.
-  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  + scale_fill_viridis_d()
+  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7))  + scale_fill_viridis_d()
 ```
@@
 #| label: fig-shootings-map
-#| fig-cap: Location of police facilities in relation to Toronto neighbourhoods with the highest number of shootings from 2018 to 2023.
+#| fig-cap: Location of police facilities in relation to six Toronto neighbourhoods with the highest number of shootings from 2018 to 2023.
 #| echo: false

### 2024-09-25 paper/paper.qmd
Added in ward profile table

@@
 ## Additional Tables 
+```{r}
+#| label: tbl-wardprofile
+#| tbl-cap:  Average total income of households in 2020 across the 25 wards of Toronto.
+#| echo: false
+#| warning: false
+#| message: false
+
+# Creates table
+cleaned_ward_data |>
+  rename(
+    "Ward Number" = ward_num,
+    "Ward Name" = ward_name,
+    "Average total income of households in 2020 ($)" = avg_income
+  ) |>
+  kable()
+```
 

### 2024-09-27 paper/paper.qmd
Finished simulation script and updated paper

@@
 # Results {#sec-results}
-## Number of Homicide and Shooting Cases in Toronto from 2018 to 2023 in Toronto
 ```{r}
@@
   cleaned_crime_data |>
+  unique() |>
   # Calculate the total sum of cases across all neighbourhoods 
   mutate(
-    num_cases_total = sum(num_of_cases)
+    num_cases_total = sum(num_of_cases),
+    population_of_toronto = sum(population_2023)
   ) |>
-  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total) |>
+  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total, population_2023, population_of_toronto) |>
   # Groups data by type of crime, and neighbourhood
@@
     num_cases_total_per_neighbourhood_per_crime = sum(num_of_cases),
-    num_cases_total_per_neighbourhood_per_crime_proportion = round(num_cases_total_per_neighbourhood_per_crime / num_cases_total, 3)
+    num_cases_total_per_neighbourhood_per_crime_proportion = round(num_cases_total_per_neighbourhood_per_crime / num_cases_total, 2)
   ) |>
@@
     num_cases_total_per_neighbourhood = sum(num_of_cases),
-    num_cases_total_per_neighbourhood_proportion = round(num_cases_total_per_neighbourhood / num_cases_total, 3)
+    num_cases_total_per_neighbourhood_proportion = round(num_cases_total_per_neighbourhood / num_cases_total, 2),
+    proportion_of_toronto_pop_in_neighbourhood = round(population_2023 / population_of_toronto, 2)
   ) |>
+  ungroup() |>
   # Obtains unique rows
@@
   tibble() |>
-  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion)
+  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion, proportion_of_toronto_pop_in_neighbourhood)
 
@@
   cleaned_crime_data |>
+  unique() |>
   # Calculate the total sum of cases across all neighbourhoods 
@@
     yearly_num_cases_total = sum(num_of_cases),
-    yearly_num_cases_total_proportion = round(yearly_num_cases_total / num_cases_total, 3)
+    yearly_num_cases_total_proportion = round(yearly_num_cases_total / num_cases_total, 2)
   ) |>
@@
 ```
+## Number of Homicide and Shooting Cases in Toronto from 2018 to 2023 in Toronto
 
+Looking at @tbl-totalcasestable and @fig-totalcaseshistogram, we can see that the number of shootings across all Toronto neighbourhoods increased from 427 cases in 2018 to 492 cases in 2019. 
 ```{r}
@@
   theme_minimal() + 
-  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7)) + scale_fill_viridis_d()
+  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7)) + scale_fill_brewer(palette = "Paired")
 ```
@@
   arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
-  select(neighbourhood, hood_id, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
+  select(neighbourhood, hood_id, proportion_of_toronto_pop_in_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
   unique() |>
@@
     "Neighbourhood ID" = hood_id,
-    "Neighbourhood" = neighbourhood
+    "Neighbourhood" = neighbourhood,
+    "Proportion of Population (2023)" = proportion_of_toronto_pop_in_neighbourhood
   ) |>
@@
     "Neighbourhood ID" = hood_id, 
-    "Neighbourhood" = neighbourhood
+    "Neighbourhood" = neighbourhood, 
+    "Proportion of Population (2023)" = proportion_of_toronto_pop_in_neighbourhood
   ) |>
@@
   arrange(desc(num_cases_total_per_neighbourhood)) |>
-  select(neighbourhood, hood_id, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
+  select(neighbourhood, hood_id, proportion_of_toronto_pop_in_neighbourhood, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
   unique() |>
@@
     "Neighbourhood ID" = hood_id,
-    "Neighbourhood" = neighbourhood
+    "Neighbourhood" = neighbourhood, 
+    "Proportion of Population (2023)" = proportion_of_toronto_pop_in_neighbourhood
   ) |>
@@
   # Reference https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2 to figure out how to rotate text along x axis.
-  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7))  + scale_fill_viridis_d()
+  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7))  + scale_fill_brewer(palette = "Paired")
 ```
@@
     size = 2, 
-    color = "white"
+    color = "black"
     ) +
@@
     aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
-  ) + labs(color = "Location")
+  ) + labs(color = "Location") + scale_color_manual(values = c("Police Facility" = "darkred")) + 
+  scale_fill_gradient(low = 'lightblue', high = 'darkblue')
 

### 2024-09-27 scripts/00-simulate_data.R

@@ -1,19 +1,99 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates data sets pertaining to crimes and police locations
+# Author: Emily Su
+# Date: 26 September 2024
+# Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: Have ran 00-install_packages.R prior to install
+# necessary packages.
+# Any other information needed?
+# NOTE: This script was checked through lintr for styling
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+set.seed(646)
+
+# Global variables
+population_of_toronto <- 2930000
+num_of_neighbourhoods <- 158
+
+# Dataset 1: Crime dataset
+# Expected Columns: year | neighbourhood | hood_id | num_shootings
+# (more columns) | num_homicides | population
+
+simulated_dataset_1 <- tibble(
+  # Create a vector for the years We will create a vector for the years 2018
+  # to 2023 and repeat this 158 times (to represent each neighbourhood)
+  year = rep(c(1:6 + 2017), num_of_neighbourhoods),
+  # Generate 158 unique neighbourhood names
+  neighbourhood =
+    rep(paste(rep("neighbourhood", size = num_of_neighbourhoods),
+              sample(x = 1:num_of_neighbourhoods, size = num_of_neighbourhoods,
+                     replace = FALSE), sep = " "), 6),
+  # Generate hood_id
+  hood_id =
+    as.numeric(rep(sample(x = 1:num_of_neighbourhoods,
+                          size = num_of_neighbourhoods,
+                          replace = FALSE), 6)),
+  # Create 6 * num_of_neighbourhoods random values from uniform distribution
+  # to get number of shootings each year per neighbourhood
+  num_shootings = round(runif(n = 6 * num_of_neighbourhoods, min = 0,
+                              max = 100)),
+  # Randomize number of homicides
+  num_homicides = round(runif(n = 6 * num_of_neighbourhoods,
[…]

### 2024-09-27 scripts/00-simulate_data.R
Removed unused code in scripts

@@ -6,7 +6,7 @@
 # License: MIT
 # Pre-requisites: Have ran 00-install_packages.R prior to install
 # necessary packages.
-# Any other information needed?
+# Any other information needed? -
 # NOTE: This script was checked through lintr for styling
 
 #### Workspace setup ####

### 2024-09-27 scripts/01-download_data.R

@@ -29,13 +29,6 @@ library(sf)
 # Code to download data was adapted from:
 # https://open.toronto.ca/dataset/neighbourhood-crime-rates/
 
-# # Function to download data
-# download_csv <- function(package_num) {
-#   files <- filter(list_package_resources(package_num),
-#                   tolower(format) %in% c("csv"))
-#   return(filter(files, row_number() == 1) %>% get_resource())
-# }
-
 # Function to download data
 # Downloading geojson file due to the formatting of "geometry" column
 download_geojson <- function(package_num) {

### 2024-09-27 scripts/02-data_cleaning.R

@@ -24,10 +24,6 @@ raw_crime_rates_data <-
   geojson_sf("data/raw_data/raw_crime_rates.geojson")
 raw_police_location_data <-
   geojson_sf("data/raw_data/raw_police_location.geojson")
-# raw_crime_rates_data <-
-#   read_csv("data/raw_data/raw_crime_rates.csv", show_col_types = FALSE)
-#raw_police_location_data <-
-#  read_csv("data/raw_data/raw_police_location.csv", show_col_types = FALSE)
 
 # Dataset 1 (Ward Profiles Dataset)
 # Expected columns: ward_num | ward_name | avg_income
@@ -214,10 +210,3 @@ write_csv(cleaned_ward_data, "data/analysis_data/cleaned_ward_data.csv")
 # Save police location data as a geojson
 write_sf(cleaned_police_location,
           "data/analysis_data/cleaned_police_location.geojson")
-# # Save crime data as a CSV
-# write_csv(cleaned_crime_data, "data/analysis_data/cleaned_crime_data.csv")
-# # Save ward profile data as a CSV
-# write_csv(cleaned_ward_data, "data/analysis_data/cleaned_ward_data.csv")
-# # Save police location data as a CSV
-# write_csv(cleaned_police_location,
-#           "data/analysis_data/cleaned_police_location.csv")

### 2024-09-27 scripts/00-simulate_data.R
Finished test data script and fixed simulation data script

@@ -17,7 +17,7 @@ set.seed(646)
 
 # Global variables
 population_of_toronto <- 2930000
-num_of_neighbourhoods <- 158
+num_of_neighbourhoods <- 174
 
 # Dataset 1: Crime dataset
 # Expected Columns: year | neighbourhood | hood_id | num_shootings
@@ -59,10 +59,10 @@ simulated_dataset_1 <- tibble(
 simulated_dataset_1$year |> min() == 2018
 simulated_dataset_1$year |> max() == 2023
 
-# 2. Made sure "hood_id" has neighbourhood id that is from 1 to 158
-simulated_dataset_1$hood_id |> unique() %in% c(1:158)
+# 2. Made sure "hood_id" has neighbourhood id that is from 1 to 174
+simulated_dataset_1$hood_id |> unique() %in% c(1:174)
 simulated_dataset_1$hood_id |> min() == 1
-simulated_dataset_1$hood_id |> max() == 158
+simulated_dataset_1$hood_id |> max() == 174
 
 # 3. "num_shootings" has min greater than or equal to 0
 simulated_dataset_1$num_shootings |> min() >= 0

### 2024-09-27 scripts/03-test_data.R

@@ -1,15 +1,63 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests cleaned datasets about crimes, police locations,
+# and ward data. 
+# Author: Emily Su
+# Date: 26 September 2024
+# Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: Have ran 00-install_packages.R, 01-download_data.R,
+# and 02-data_cleaning.R to obtain cleaned datasets. 
+# Any other information needed? -
+# NOTE: This script was checked through lintr for styling
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(geojsonsf)
+
+#### Read in data ####
+cleaned_ward_data <-
+  read_csv("data/analysis_data/cleaned_ward_data.csv", show_col_types = FALSE)
+cleaned_crime_data <-
+  geojson_sf("data/analysis_data/cleaned_crime_data.geojson")
+cleaned_police_location <-
+  geojson_sf("data/analysis_data/cleaned_police_location.geojson")
 
 #### Test data ####
+
+# Crime dataset: Data Validation / Tests
+# 1. "year" contains years between 2018 to 2023
+cleaned_crime_data$year |> min() == 2018
+cleaned_crime_data$year |> max() == 2023
+
+# 2. Made sure "hood_id" has neighbourhood id that is from 1 to 174
+cleaned_crime_data$hood_id |> unique() %in% c(1:174)
+cleaned_crime_data$hood_id |> min() == 1
+cleaned_crime_data$hood_id |> max() == 174
+
+# 3. "num_of_cases" has min greater than or equal to 0
+cleaned_crime_data$num_of_cases |> min() >= 0
+
+# 4. "population_2023" has min greater than or equal to 0
+cleaned_crime_data$population |> min() >= 0
+
+# 5. Check the classes
+class(cleaned_crime_data$year) == "numeric"
+class(cleaned_crime_data$num_of_cases) == "numeric"
+class(cleaned_crime_data$hood_id) == "numeric"
+class(cleaned_crime_data$homicide_or_shooting) == "character"
[…]

### 2024-09-27 scripts/00-install_packages.R
Fixed scripts' styling

@@ -17,5 +17,5 @@ install.packages("dplyr")
 install.packages("ggplot2") # To make graphs
 install.packages("lintr") # To check styling of code
 install.packages("styler") # To style code
-install.packages("geojsonsf") 
+install.packages("geojsonsf")
 install.packages("sf")

### 2024-09-27 scripts/01-download_data.R

@@ -38,7 +38,6 @@ download_geojson <- function(package_num) {
 }
 
 # Download "Neighbourhood Crime Rates" Dataset
-# crime_rates <- download_csv("neighbourhood-crime-rates")
 crime_rates <- download_geojson("neighbourhood-crime-rates")
 
 # Download "Police Facility Locations" Dataset
@@ -50,7 +49,6 @@ ward_profiles <-
          row_number() == 1) |> get_resource()
 
 #### Save data ####
-# write_csv(crime_rates, "data/raw_data/raw_crime_rates.csv")
 write_sf(crime_rates, "data/raw_data/raw_crime_rates.geojson")
 write_sf(police_location, "data/raw_data/raw_police_location.geojson")
 write_csv(as.data.frame(ward_profiles[1]),

### 2024-09-27 scripts/02-data_cleaning.R

@@ -33,49 +33,49 @@ cleaned_ward_data <- raw_ward_profiles_data |> clean_names()
 # Get row that contains the average household income
 cleaned_ward_data <-
   cleaned_ward_data |>
-  filter(x2021_one_variable_city_of_toronto_profiles == 
+  filter(x2021_one_variable_city_of_toronto_profiles ==
            "Average total income of households in 2020 ($)")
-# Convert row to a column 
-cleaned_ward_data <- 
+# Convert row to a column
+cleaned_ward_data <-
   cleaned_ward_data |>
-  # Code snippet to convert to a 1-column matrix and then to a vector 
+  # Code snippet to convert to a 1-column matrix and then to a vector
   # was adapted from the following:
   # https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
   t() |> # Convert 1-row matrix into a 1-column matrix
   c() |> # Convert to a vector
   tibble() |> # Change into a tibble
-  slice(-1) |> # Remove the first row 
+  slice(-1) |> # Remove the first row
   slice(-1) |>
   rename( # Left side is the new name and the right side is the old name
     avg_income = "c(t(cleaned_ward_data))"
-  ) |> 
+  ) |>
   mutate(
-    ward_num = c(1:25), 
-    ward_name = c("Etobicoke North", 
-                      "Etobicoke Centre", 
-                      "Etobicoke-Lakeshore", 
-                      "Parkdale-High Park", 
-                      "York South-Weston", 
-                      "York Centre", 
-                      "Humber River-Black Creek",
-                      "Eglinton-Lawrence", 
-                      "Davenport", 
-                      "Spadina-Fort York", 
-                      "University-Rosedale", 
-                      "Toronto-St. Paul’s", 
-                      "Toronto Centre", 
-                      "Toronto-Danforth", 
-                      "Don Valley West", 
-                      "Don Valley East",
-                      "Don Valley North", 
-                      "Willowdale", 
-                      "Beaches-East York", 
-                      "Scarborough Southwest", 
-                      "Scarborough Centre",
-                      "Scarborough-Agincourt",
-                      "Scarborough North", 
[…]

### 2024-09-27 scripts/03-test_data.R

@@ -1,12 +1,12 @@
 #### Preamble ####
 # Purpose: Tests cleaned datasets about crimes, police locations,
-# and ward data. 
+# and ward data.
 # Author: Emily Su
 # Date: 26 September 2024
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Have ran 00-install_packages.R, 01-download_data.R,
-# and 02-data_cleaning.R to obtain cleaned datasets. 
+# and 02-data_cleaning.R to obtain cleaned datasets.
 # Any other information needed? -
 # NOTE: This script was checked through lintr for styling
 
@@ -55,9 +55,9 @@ class(cleaned_police_location$facility) == "character"
 # cleaned_ward_data: Data Validation / Tests
 # 1. Check if that are ward numbers from 1 to 25
 cleaned_ward_data$ward_num |> unique() %in% c(1:25)
-# 2. Check that average income is equal to or greater than 0 
+# 2. Check that average income is equal to or greater than 0
 cleaned_ward_data$avg_income |> min() >= 0
-# 3. Check classes 
+# 3. Check classes
 class(cleaned_ward_data$ward_num) == "numeric"
 class(cleaned_ward_data$ward_name) == "character"
 class(cleaned_ward_data$avg_income) == "numeric"

### 2024-09-27 scripts/02-data_cleaning.R
Updated data cleaning script and references

@@ -202,7 +202,8 @@ cleaned_crime_data <-
   # rename
   rename(
     "neighbourhood" = area_name
-  )
+  ) |>
+  unique()
 
 # Dataset 3
 # Expected Columns: facility | geometry
@@ -211,7 +212,7 @@ cleaned_police_location <- raw_police_location_data |> clean_names()
 
 # Select expected columns
 cleaned_police_location <-
-  cleaned_police_location |> select(facility, geometry)
+  cleaned_police_location |> select(facility, geometry) |> unique()
 
 #### Save data ####
 # Save crime data as a geojson

### 2024-09-27 scripts/00-simulate_data.R
Finish simulation and testing script

@@ -17,7 +17,8 @@ set.seed(646)
 
 # Global variables
 population_of_toronto <- 2930000
-num_of_neighbourhoods <- 174
+num_of_neighbourhood_ids <- 174
+num_of_neighbourhoods <- 158
 
 # Dataset 1: Crime dataset
 # Expected Columns: year | neighbourhood | hood_id | num_shootings
@@ -34,7 +35,7 @@ simulated_dataset_1 <- tibble(
                      replace = FALSE), sep = " "), 6),
   # Generate hood_id
   hood_id =
-    as.numeric(rep(sample(x = 1:num_of_neighbourhoods,
+    as.numeric(rep(sample(x = 1:num_of_neighbourhood_ids,
                           size = num_of_neighbourhoods,
                           replace = FALSE), 6)),
   # Create 6 * num_of_neighbourhoods random values from uniform distribution
@@ -63,6 +64,9 @@ simulated_dataset_1$year |> max() == 2023
 simulated_dataset_1$hood_id |> unique() %in% c(1:174)
 simulated_dataset_1$hood_id |> min() == 1
 simulated_dataset_1$hood_id |> max() == 174
+
+# 2. Check that there are 158 neighbourhoods
+simulated_dataset_1$neighbourhood |> n_distinct() == 158
 
 # 3. "num_shootings" has min greater than or equal to 0
 simulated_dataset_1$num_shootings |> min() >= 0

### 2024-09-27 scripts/03-test_data.R

@@ -13,6 +13,7 @@
 #### Workspace setup ####
 library(tidyverse)
 library(geojsonsf)
+library(dplyr)
 
 #### Read in data ####
 cleaned_ward_data <-
@@ -34,13 +35,16 @@ cleaned_crime_data$hood_id |> unique() %in% c(1:174)
 cleaned_crime_data$hood_id |> min() == 1
 cleaned_crime_data$hood_id |> max() == 174
 
-# 3. "num_of_cases" has min greater than or equal to 0
+# 3. Check that there are 158 neighbourhoods
+cleaned_crime_data$neighbourhood |> n_distinct() == 158
+
+# 4. "num_of_cases" has min greater than or equal to 0
 cleaned_crime_data$num_of_cases |> min() >= 0
 
-# 4. "population_2023" has min greater than or equal to 0
+# 5. "population_2023" has min greater than or equal to 0
 cleaned_crime_data$population |> min() >= 0
 
-# 5. Check the classes
+# 6. Check the classes
 class(cleaned_crime_data$year) == "numeric"
 class(cleaned_crime_data$num_of_cases) == "numeric"
 class(cleaned_crime_data$hood_id) == "numeric"

### 2024-09-28 scripts/01-download_data.R
Deleted ward profile dataset

@@ -1,7 +1,6 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from Open Data Toronto on the location
-# of police facilities, neighbourhood crime rates, and the profiles of
-# Toronto's 25 wards.
+# of police facilities and neighbourhood crime rates.
 # Author: Emily Su
 # Date: 21 September 2024
 # Contact: em.su@mail.utoronto.ca
@@ -12,8 +11,6 @@
 # NOTE: This script was checked through lintr for styling
 
 # Datasets to be downloaded and saved
-# Ward Profiles (25-Ward Model):
-# - https://open.toronto.ca/dataset/ward-profiles-25-ward-model/
 # Police Facility Location
 # - https://open.toronto.ca/dataset/police-facility-locations/
 # Neighbourhood Crime Rates
@@ -43,13 +40,7 @@ crime_rates <- download_geojson("neighbourhood-crime-rates")
 # Download "Police Facility Locations" Dataset
 police_location <- download_geojson("9aeefa17-27e8-4dd9-b74d-80f7f9eb85ac")
 
-# Download "Ward Profiles (25-Ward Model)"
-ward_profiles <-
-  filter(list_package_resources("6678e1a6-d25f-4dff-b2b7-aa8f042bc2eb"),
-         row_number() == 1) |> get_resource()
 
 #### Save data ####
 write_sf(crime_rates, "data/raw_data/raw_crime_rates.geojson")
 write_sf(police_location, "data/raw_data/raw_police_location.geojson")
-write_csv(as.data.frame(ward_profiles[1]),
-          "data/raw_data/raw_ward_profiles.csv")

### 2024-09-28 scripts/02-data_cleaning.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the unedited ward profile data, unedited police location
+# Purpose: Cleans the unedited police location
 # data, and unedited data on neighbourhood crime rates.
 # Author: Emily Su
 # Date: 21 September 2024
@@ -18,75 +18,12 @@ library(geojsonsf)
 library(sf)
 
 #### Clean data ####
-raw_ward_profiles_data <-
-  read_csv("data/raw_data/raw_ward_profiles.csv", show_col_types = FALSE)
 raw_crime_rates_data <-
   geojson_sf("data/raw_data/raw_crime_rates.geojson")
 raw_police_location_data <-
   geojson_sf("data/raw_data/raw_police_location.geojson")
 
-# Dataset 1 (Ward Profiles Dataset)
-# Expected columns: ward_num | ward_name | avg_income
-# Clean column names
-cleaned_ward_data <- raw_ward_profiles_data |> clean_names()
-
-# Get row that contains the average household income
-cleaned_ward_data <-
-  cleaned_ward_data |>
-  filter(x2021_one_variable_city_of_toronto_profiles ==
-           "Average total income of households in 2020 ($)")
-# Convert row to a column
-cleaned_ward_data <-
-  cleaned_ward_data |>
-  # Code snippet to convert to a 1-column matrix and then to a vector
-  # was adapted from the following:
-  # https://www.geeksforgeeks.org/convert-matrix-to-vector-in-r/
-  t() |> # Convert 1-row matrix into a 1-column matrix
-  c() |> # Convert to a vector
-  tibble() |> # Change into a tibble
-  slice(-1) |> # Remove the first row
-  slice(-1) |>
-  rename( # Left side is the new name and the right side is the old name
-    avg_income = "c(t(cleaned_ward_data))"
-  ) |>
-  mutate(
-    ward_num = c(1:25),
-    ward_name = c("Etobicoke North",
-                  "Etobicoke Centre",
-                  "Etobicoke-Lakeshore",
-                  "Parkdale-High Park",
-                  "York South-Weston",
-                  "York Centre",
-                  "Humber River-Black Creek",
-                  "Eglinton-Lawrence",
-                  "Davenport",
-                  "Spadina-Fort York",
-                  "University-Rosedale",
-                  "Toronto-St. Paul’s",
-                  "Toronto Centre",
-                  "Toronto-Danforth",
-                  "Don Valley West",
[…]

### 2024-09-28 scripts/03-test_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
-# Purpose: Tests cleaned datasets about crimes, police locations,
-# and ward data.
+# Purpose: Tests cleaned datasets about crimes
+# and police locations data.
 # Author: Emily Su
 # Date: 26 September 2024
 # Contact: em.su@mail.utoronto.ca
@@ -16,8 +16,6 @@ library(geojsonsf)
 library(dplyr)
 
 #### Read in data ####
-cleaned_ward_data <-
-  read_csv("data/analysis_data/cleaned_ward_data.csv", show_col_types = FALSE)
 cleaned_crime_data <-
   geojson_sf("data/analysis_data/cleaned_crime_data.geojson")
 cleaned_police_location <-
@@ -55,13 +53,3 @@ class(cleaned_crime_data$population_2023) == "numeric"
 # Police location dataset: Data Validation / Tests
 # Check the classes
 class(cleaned_police_location$facility) == "character"
-
-# cleaned_ward_data: Data Validation / Tests
-# 1. Check if that are ward numbers from 1 to 25
-cleaned_ward_data$ward_num |> unique() %in% c(1:25)
-# 2. Check that average income is equal to or greater than 0
-cleaned_ward_data$avg_income |> min() >= 0
-# 3. Check classes
-class(cleaned_ward_data$ward_num) == "numeric"
-class(cleaned_ward_data$ward_name) == "character"
-class(cleaned_ward_data$avg_income) == "numeric"

### 2024-09-28 paper/paper.qmd
Updated paper

@@
 # Police data 
-cleaned_police_location <- geojson_sf('../data/analysis_data/cleaned_police_location.geojson')
+cleaned_police_location <- geojson_sf('../data/analysis_data/cleaned_police_location.geojson') |> unique()
 # Crime data
-cleaned_crime_data <- geojson_sf('../data/analysis_data/cleaned_crime_data.geojson')
-# Ward data 
-cleaned_ward_data <- read_csv('../data/analysis_data/cleaned_ward_data.csv', show_col_types = FALSE)
+cleaned_crime_data <- geojson_sf('../data/analysis_data/cleaned_crime_data.geojson') |> unique()
 ```
@@
 # Introduction
-
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
 
@@
 ## Methodology
+The three datasets used by the paper were simulated, retrieved, cleaned, analyzed, and tested 
+using the R programming language [@citeR], tidyverse [@citetidyverse], opendatatoronto [@citeopendatatoronto], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], GeoJSON [@citegeojsonsf], and sf [@citesf].
 
 ## Data Source and Measurements
+The data used for the analysis comes from the "cleaned_crime_data" and "cleaned_police_location" datasets.
+The datasets was constructed by calling the opendatatoronto package [@citeopendatatoronto], which accesses and download data from the City of Toronto's Open Data Portal.
+The latest versions of the datasets was scraped on September 24, 2024 at 10:28 pm EDT.
+We also removed duplicate observations from the original dataset that comes from reading in GeoJSON files for the datasets.
+Other datasets we considered scraping includes the "Neighbourhoods" dataset to obtain neighbourhood boundaries for the maps but further testing showed that these boundaries were provided by the "cleaned_crime_data" dataset.
 
+### Neighbourhood Crime Rates Dataset
+The "cleaned_crime_data" dataset was retrieved from the "Neighbourhood Crime Rates" dataset and the last time the dataset was updated on the Open Data Portal was on January 11, 2024.
+This dataset comprised of crime data by neighbourhood that included the number of cases of assault, robbery, homicide, shootings, auto theft, break and enter, and theft over.
+The original dataset was published by the Toronto Police Services collected from 2014 to 2023, inclusive and is updated annually.
+For the cleaned dataset, we decided to filter for homicides and shootings and focus on the cases from 2018 to 2023, inclusive.
+The 2023 population projection in the dataset was provided by Environics Analytics where they counted only the residents in the neighbourhood [@citeneighbourhoodcrimerates].
+
+Each observation in the dataset represent one of Toronto's 158 neighbourhoods and its shootings and homicides from the years 2018 to 2023, inclusive that was recorded by the Toronto Police Services.
+There are 1896 unique observations in total in the dataset.
+
+Looking at @tbl-dataset-crime-preview, year represents the year the number of homicides and shootings takes place in the neighbourhood and num_of_cases and homicide_or_shooting were constructed based on the columns representing homicide and shootings in the original dataset.
+homicide_or_shooting indicates what the current crime in the neighbourhood the observation is describing and the num_of_cases indicates the number of cases for the crime indicated by homicide_or_shooting.
+hood_id is a number between 1 to 174 that represents the ID for the neighbourhood with some neighbourhoods having more than 1 ID.
+The neighbourhood column contains the name of the neighbourhood.
+
+Not shown in @tbl-dataset-crime-preview, the geometry column contains points to create the neighbourhood area the shooting or homicide takes place in and population_2023 represents the population projection for 2023 obtained from Environics Analytics.
+Other variables constructed from this dataset that is not shown includes num_cases_total, population_of_toronto, num_cases_total_per_neigh_per_crime, num_cases_total_per_neigh_per_crime_prop, num_cases_total_per_neigh, num_cases_total_per_neigh_prop, prop_of_toronto_pop_in_neigh, yearly_num_cases_total, and yearly_num_cases_total_prop.
+num_cases_total is the total number of shootings and homicides from 2018 to 2023 while population_of_toronto is the population of Toronto in 2023 obtained from summing up the population_2023 column.
+num_cases_total_per_neigh_per_crime represents the number of cases for either homicides or shootings in a particular neighbourhood while num_cases_total_per_neigh_per_crime_prop is the proportion of num_cases_total_per_neigh_per_crime in relation to num_cases_total.
+num_cases_total_per_neigh represents the number of cases total for each neighbourhood with num_cases_total_per_neigh_prop representing the proportion of num_cases_total_per_neigh in relation to num_cases_total.
+prop_of_toronto_pop_in_neigh is the population of the neighbourhood in relation to the population of Toronto in 2023.
+yearly_num_cases_total represents the total number of cases for each crime yearly across ALL neighbourhood per year and yearly_num_cases_total_prop is the proportion of yearly_num_cases_total in relation to num_cases_total.
+
+```{r}
+#| label: tbl-dataset-crime-preview
+#| tbl-cap: Preview of dataset on crime rates across Toronto neighbourhoods from 2018 to 2023 provided by Toronto Police Services
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create table displaying data
+cleaned_crime_data |>
+  tibble() |>
+  select(year, num_of_cases, hood_id, homicide_or_shooting, neighbourhood) |>
+  head() |>
+  kable()
+```
+### Police Facility Locations Dataset
+The "cleaned_police_location" dataset was obtained from the "Police Facility Locations" dataset on the Open Data Portal.
+The last time the dataset was updated on January 20, 2023 and is updated as new locations are added.
+Each observation in the dataset represents a police facility in Toronto and its respective geographical point that is recorded by Toronto Police Services as new locations open.
+There are 26 unique observations in the dataset and contains two variables of interest facility and geometry.
+As seen in @tbl-dataset-police-preview, facility represents the name of the police location while geometry represents a geographical point of the location on a map.
+
+```{r}
+#| label: tbl-dataset-police-preview
+#| tbl-cap: Preview of dataset on police facility locations provided by Toronto Police Services
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create table displaying data
+cleaned_police_location |>
+  head() |>
+  kable()
+```
 ## Variables of Interest
+Our variables of interest from the "cleaned_crime_data" dataset are num_of_cases, homicide_or_shooting, hood_id, neighbourhood, and geometry.
+The variables constructed based on the dataset that are variables of interest includes num_cases_total, num_cases_total_per_neigh_per_crime, num_cases_total_per_neigh_per_crime_prop, num_cases_total_per_neigh, num_cases_total_per_neigh_prop, prop_of_toronto_pop_in_neigh, yearly_num_cases_total, and yearly_num_cases_total_prop.
+The variables of interest from the "cleaned_police_location" dataset includes facility and geometry.
 
@@
   mutate (
-    num_cases_total_per_neighbourhood_per_crime = sum(num_of_cases),
-    num_cases_total_per_neighbourhood_per_crime_proportion = round(num_cases_total_per_neighbourhood_per_crime / num_cases_total, 2)
+    num_cases_total_per_neigh_per_crime = sum(num_of_cases),
+    num_cases_total_per_neigh_per_crime_prop = round(num_cases_total_per_neigh_per_crime / num_cases_total, 2)
   ) |>
@@
   mutate(
-    num_cases_total_per_neighbourhood = sum(num_of_cases),
-    num_cases_total_per_neighbourhood_proportion = round(num_cases_total_per_neighbourhood / num_cases_total, 2),
-    proportion_of_toronto_pop_in_neighbourhood = round(population_2023 / population_of_toronto, 2)
+    num_cases_total_per_neigh = sum(num_of_cases),
+    num_cases_total_per_neigh_prop = round(num_cases_total_per_neigh / num_cases_total, 2),
+    prop_of_toronto_pop_in_neigh = round(population_2023 / population_of_toronto, 2)
   ) |>
@@
   # Arrange based on the total number of cases per neighbourhood
-  arrange(desc(num_cases_total_per_neighbourhood)) |>
+  arrange(desc(num_cases_total_per_neigh)) |>
   tibble() |>
-  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion, proportion_of_toronto_pop_in_neighbourhood)
+  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neigh, num_cases_total_per_neigh_per_crime, num_cases_total_per_neigh_prop, num_cases_total_per_neigh_per_crime_prop, prop_of_toronto_pop_in_neigh)
 
@@
     yearly_num_cases_total = sum(num_of_cases),
-    yearly_num_cases_total_proportion = round(yearly_num_cases_total / num_cases_total, 2)
+    yearly_num_cases_total_prop = round(yearly_num_cases_total / num_cases_total, 2)
   ) |>
   tibble() |>
-  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_proportion) |>
+  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_prop) |>
   unique() |>
@@
 Looking at @tbl-totalcasestable and @fig-totalcaseshistogram, we can see that the number of shootings across all Toronto neighbourhoods increased from 427 cases in 2018 to 492 cases in 2019.
+In 2020, the number of shootings decreased to 462 and it decreased the following years for 2021 with 409 shootings, for 2022 with 380 shootings, and for 2023 with 342 shootings.
+For homicides, there were 98 cases in 2018, 79 cases in 2019, and 71 cases in 2020.
+There was an increase in homicides in 2021 to 85 cases and decreases in homicides in 2022 and 2023 to 71 and 72 cases, respectively.
 ```{r}
@@
 yearly_crime |>
-  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_proportion) |>
+  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_prop) |>
   # Arrange based on year
@@
     "Total number of cases yearly" = yearly_num_cases_total,
-    "Proportion of cases yearly" = yearly_num_cases_total_proportion,
+    "Proportion of cases yearly" = yearly_num_cases_total_prop,
     "Year" = year
@@
 
-```{r}
-#| label: tbl-totalcasesstats
-#| tbl-cap: Minimum, quartiles, median, and maximum of the number of homicides and shootings in Toronto from 2018 to 2023
-#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
-#| warning: false # Warnings are not included 
-#| message: false # Messages are not included
-
-# Creates table
-yearly_crime |>
-  select(yearly_num_cases_total) |>
-  rename(
-    "Total number of cases" = yearly_num_cases_total
-  ) |>
-  summary() |>
-  kable()
-```
+Across all Toronto neighbourhoods, @tbl-totalshootingsstats shows that the mean total number of shootings is about 419 cases from 2018 to 2023 across all neighbourhoods yearly.
+The median total number of shootings is about 418 cases from 2018 to 2023 across all neighbourhoods yearly.
+@tbl-totalhomicidesstats shows that the mean number of homicides annually across all Toronto neighbourhoods from 2018 to 2023 is 79 while the median is 76 cases.
 
@@
 
-## Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings
+## Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings from 2018 to 2023
 
+@tbl-highestcasesshootingstable shows that Glenfield-Jane Heights has the highest number of shootings from 2018 to 2023 with 133 cases (4% of all shootings and homicides from 2018 to 2023) out of the 6 Toronto neighbourhoods with the highest number of shootings.
+Following this, Mount Olive-Silverstone-Jamestown has second highest number of shootings with 82 cases (3% of all shootings and homicides from 2018 to 2023).
+For the rest in descending order, Black Creek has 69 cases (2% of all shootings and homicides), York University Heights has 59 cases (2% of all shootings and homicides), Yorkdale-Glen Park has 58 cases (2% of all shootings and homicides), and Golfdale-Cedarbrae-Woburn has 54 cases (2% of all shootings and homicides).
+
+@tbl-highestcaseshomicidestable indicates that Moss Park has the highest number of homicides from 2018 to 2023 with 20 cases (1% of all shootings and homicides).
+Mount Olive-Silverstone-Jamestown is not only in second for the highest number of shootings but also for homicides with 15 cases from 2018 to 2023 (1% of all shootings and homicides).
+For the other four neighbourhoods, Avondale has 13 cases ($<1%$ of all shootings and homicides), Glenfield-Jane Heights and York University Heights both have 11 cases ($<1%$ of all shootings and homicides), and Eglinton East	has 10 cases ($<1%$ of all shootings and homicides).
+
+For both @tbl-highestcasesshootingstable and @tbl-highestcaseshomicidestable, the proportion of people living in each of the 12 neighbourhoods are $<1$. 
 ```{r}
@@
   filter(homicide_or_shooting == "shooting") |>
-  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
-  select(neighbourhood, hood_id, proportion_of_toronto_pop_in_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
+  arrange(desc(num_cases_total_per_neigh_per_crime)) |>
+  select(neighbourhood, hood_id, prop_of_toronto_pop_in_neigh, num_cases_total_per_neigh_per_crime, num_cases_total_per_neigh_per_crime_prop) |>
   unique() |>
   rename(
-    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
-    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
+    "Total number of cases (2018 to 2023)" = num_cases_total_per_neigh_per_crime,
+    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neigh_per_crime_prop,
     "Neighbourhood ID" = hood_id,
     "Neighbourhood" = neighbourhood,
-    "Proportion of Population (2023)" = proportion_of_toronto_pop_in_neighbourhood
+    "Proportion of Population (2023)" = prop_of_toronto_pop_in_neigh
   ) |>
@@
   filter(homicide_or_shooting == "homicide") |>
-  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
-  select(neighbourhood, hood_id, proportion_of_toronto_pop_in_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
+  arrange(desc(num_cases_total_per_neigh_per_crime)) |>
+  select(neighbourhood, hood_id, prop_of_toronto_pop_in_neigh, num_cases_total_per_neigh_per_crime, num_cases_total_per_neigh_per_crime_prop) |>
   unique() |>
   rename(
-    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
-    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
+    "Total number of cases (2018 to 2023)" = num_cases_total_per_neigh_per_crime,
+    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neigh_per_crime_prop,
     "Neighbourhood ID" = hood_id, 
     "Neighbourhood" = neighbourhood, 
-    "Proportion of Population (2023)" = proportion_of_toronto_pop_in_neighbourhood
+    "Proportion of Population (2023)" = prop_of_toronto_pop_in_neigh
   ) |>
@@
 ```
+
+Based on @tbl-highestcasestable and @fig-highestcaseshistogram, Glenfield-Jane Heights is the neighbourhood with the highest number of homicides and shootings from 2018 to 2023 with 144 cases (5% of all shootings and homicides).
+Mount Olive-Silverstone-Jamestown has the second highest number of cases with 97 cases (3% of all shootings and homicides).
+For the rest of four neighbourhoods, Black Creek has 77 cases (3% of shootings and homicides), York University Heights has 70 cases (2% of shootings and homicides), Yorkdale-Glen Park has 64 cases (2% of shootings and homicides), and West Humber-Clairville has 60 cases (2% of shootings and homicides).
+For all neighbourhoods, the proportion of the people living in each of the neighbourhood is $<1$.
 
@@
   filter(homicide_or_shooting == "homicide") |>
-  arrange(desc(num_cases_total_per_neighbourhood)) |>
-  select(neighbourhood, hood_id, proportion_of_toronto_pop_in_neighbourhood, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
+  arrange(desc(num_cases_total_per_neigh)) |>
+  select(neighbourhood, hood_id, prop_of_toronto_pop_in_neigh, num_cases_total_per_neigh, num_cases_total_per_neigh_prop) |>
   unique() |>
   rename(
-    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood,
-    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_proportion,
+    "Total number of cases (2018 to 2023)" = num_cases_total_per_neigh,
+    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neigh_prop,
     "Neighbourhood ID" = hood_id,
     "Neighbourhood" = neighbourhood, 
-    "Proportion of Population (2023)" = proportion_of_toronto_pop_in_neighbourhood
+    "Proportion of Population (2023)" = prop_of_toronto_pop_in_neigh
   ) |>
@@
 crime_per_neighbourhood |> 
-  select(hood_id, homicide_or_shooting, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime) |>
-  arrange(desc(num_cases_total_per_neighbourhood)) |>
-  select(-c(num_cases_total_per_neighbourhood)) |>
+  select(hood_id, homicide_or_shooting, num_cases_total_per_neigh, num_cases_total_per_neigh_per_crime) |>
+  arrange(desc(num_cases_total_per_neigh)) |>
+  select(-c(num_cases_total_per_neigh)) |>
   unique() |>
   head(12) |>
-  ggplot(mapping = aes(x = as.character(hood_id), y = num_cases_total_per_neighbourhood_per_crime, fill = homicide_or_shooting)) +
+  ggplot(mapping = aes(x = as.character(hood_id), y = num_cases_total_per_neigh_per_crime, fill = homicide_or_shooting)) +
   geom_bar(stat="identity", position = "dodge2") +
@@
 
-## Police Facility Locations in relation to the Number of Homicides and Shootings from 2018 to 2023 in Toronto
+## Police Facility Locations in relation to the Number of Homicides and Shootings from 2018 to 2023 across Toronto Neighbourhoods
+
+@fig-homicides-map shows that there are a greater density of police locations closer to the downtown area of Toronto.
+We can see that neighbourhood 73 also known as Moss Park, which had the number of cases of homicides according to @tbl-highestcaseshomicidestable has a police location situated along its border.
+Mount Olive-Silverstone-Jamestown (neighbourhood 2) has a police station located in the neighbourhood.
+However with other neighbourhoods with a high number of homicides such as West Humber-Clairville (1), Avondale (153), Glenfield-Jane Heights (25), York University Heights (27), and Eglinton East (138) has no police facilities in their neighbourhoods.
 
@@
 #| label: fig-homicides-map
-#| fig-cap: Location of police facilities in relation to Toronto neighbourhoods with the highest number of homicides from 2018 to 2023.
+#| fig-cap: Location of police facilities in relation to the number of homicides across Toronto neighbourhoods from 2018 to 2023.
 #| echo: false
@@
 #| message: false
+
+homicide_map_data <-
+  cleaned_crime_data |>
+  unique() |>
+  filter(homicide_or_shooting == "homicide") |>
+  # Groups data by type of crime, and neighbourhood
+  group_by(neighbourhood, homicide_or_shooting) |>
+  # Create new columns for number of cases for homicides and shootings per neighbourhood. Also calculate the proportion as well. 
+  mutate (
+    num_cases_total_per_neigh_per_crime = sum(num_of_cases)
+  ) |>
+  ungroup()
+
 # Reference the following for code:
@@
   # Fill in area regarding the number of cases 
-  geom_sf(data = cleaned_crime_data |> filter(homicide_or_shooting == "homicide"), aes(fill = num_of_cases, geometry = geometry)) +
+  geom_sf(data = homicide_map_data, aes(fill = num_cases_total_per_neigh_per_crime, geometry = geometry)) +
   # Changes the font of the title and removes the x and y axis 
@@
   geom_sf_text(
-    data = cleaned_crime_data |> filter(num_of_cases >= 14),
-    aes(label = hood_id),
-    check_overlap = TRUE,
-    size = 2, 
-    color = "black"
-    ) +
-    theme(plot.title = element_text(size = 9, face = 'bold'), 
-          plot.subtitle = element_text(size = 9),
-          rect = element_blank(), 
-          axis.text.x = element_blank(), 
-          axis.text.y = element_blank(), 
-          axis.title.x = element_blank(), 
-          axis.title.y = element_blank(), 
-          axis.ticks = element_blank(), 
-          legend.title = element_text(size = 8),
-          panel.grid.major = element_blank(),
-          panel.grid.minor = element_blank()
-          ) +
-  # Add in police locations
-  geom_point(data = cleaned_police_location,
-    aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
-  ) + labs(color = "Location") + scale_color_manual(values = c("Police Facility" = "darkred")) + 
-  scale_fill_gradient(low = 'lightblue', high = 'darkblue')
-
-```
-
-```{r}
-#| label: fig-shootings-map
-#| fig-cap: Location of police facilities in relation to six Toronto neighbourhoods with the highest number of shootings from 2018 to 2023.
-#| echo: false
-#| warning: false
-#| message: false
-# Reference the following for code:
-# https://cengel.github.io/R-spatial/mapping.html
-ggplot() + 
-    # To help convert between GeoJSON to sf  https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html
-  # Fill in area regarding the number of cases 
-  geom_sf(data = cleaned_crime_data |> filter(homicide_or_shooting == "shooting"), aes(fill = num_of_cases, geometry = geometry)) +
-  # Changes the font of the title and removes the x and y axis 
-  # Reference to figured out how to add labels to maps https://tmieno2.github.io/R-as-GIS-for-Economists/create-maps.html
-  theme_minimal() + 
-  labs(
-    x = "Longitude",
-    y = "Latitude",
-    title = "Number of shootings across Toronto neighbourhoods (2018 to 2023)",
-    fill = "Number of cases"
-  ) +
-  geom_sf_text(
-    data = cleaned_crime_data |> filter(num_of_cases >= 14),
+    data = homicide_map_data,
     aes(label = hood_id),
@@
   geom_point(data = cleaned_police_location,
-    aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
+    aes(geometry = geometry, color = "Police Facility"), alpha = 0.90, stat = "sf_coordinates"
+  ) + labs(color = "Location") + scale_color_manual(values = c("Police Facility" = "darkred")) + 
+  scale_fill_gradient(low = 'lightblue', high = 'darkblue')
+
+```
+Based on @fig-shootings-map, the neighbourhood with the highest number of shootings, Glenfield-Jane Heights (25) has no police facility located in the neighbourhood.
+However, Mount Olive-Silverstone-Jamestown (2) and Black Creek (24) have police facilities located in their neighbourhoods.
+For the other three neighbourhoods, York University Heights	(27) and Golfdale-Cedarbrae-Woburn (141) have no police facilities located in the neighbourhood while Yorkdale-Glen Park (31) does.
+
+```{r}
+#| label: fig-shootings-map
+#| fig-cap: Location of police facilities in relation to the number of shootings across Toronto neighbourhoods from 2018 to 2023.
+#| echo: false
+#| warning: false
+#| message: false
+
+shootings_map_data <-
+  cleaned_crime_data |>
+  unique() |>
+  filter(homicide_or_shooting == "shooting") |>
+  # Groups data by type of crime, and neighbourhood
+  group_by(neighbourhood, homicide_or_shooting) |>
+  mutate (
+    num_cases_total_per_neigh_per_crime = sum(num_of_cases)
+  ) |>
+  ungroup()
+
+# Reference the following for code:
+# https://cengel.github.io/R-spatial/mapping.html
+ggplot() + 
+    # To help convert between GeoJSON to sf  https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html
+  # Fill in area regarding the number of cases 
+  geom_sf(data = shootings_map_data, aes(fill = num_cases_total_per_neigh_per_crime, geometry = geometry)) +
+  # Changes the font of the title and removes the x and y axis 
+  # Reference to figured out how to add labels to maps https://tmieno2.github.io/R-as-GIS-for-Economists/create-maps.html
+  theme_minimal() + 
+  labs(
+    x = "Longitude",
+    y = "Latitude",
+    title = "Number of shootings across Toronto neighbourhoods (2018 to 2023)",
+    fill = "Number of cases"
+  ) +
+  geom_sf_text(
+    data = shootings_map_data,
+    aes(label = hood_id),
+    check_overlap = TRUE,
+    size = 2, 
+    color = "white"
+    ) +
+    theme(plot.title = element_text(size = 9, face = 'bold'), 
+          plot.subtitle = element_text(size = 9),
+          rect = element_blank(), 
+          axis.text.x = element_blank(), 
+          axis.text.y = element_blank(), 
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(), 
+          axis.ticks = element_blank(), 
+          legend.title = element_text(size = 8),
+          panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank()
+          ) +
+  # Add in police locations
+  geom_point(data = cleaned_police_location,
+    aes(geometry = geometry, color = "Police Facility"), alpha = 0.90, stat = "sf_coordinates"
   ) + labs(color = "Location") + scale_color_manual(values = c("Police Facility" = "darkred")) + 
@@
 
-## Additional Figures
-
-## Additional Tables 
-```{r}
-#| label: tbl-wardprofile
-#| tbl-cap:  Average total income of households in 2020 across the 25 wards of Toronto.
-#| echo: false
-#| warning: false
-#| message: false
-
-# Creates table
-cleaned_ward_data |>
-  rename(
-    "Ward Number" = ward_num,
-    "Ward Name" = ward_name,
-    "Average total income of households in 2020 ($)" = avg_income
-  ) |>
-  kable()
-```
-
 \newpage

### 2024-09-29 paper/paper.qmd
Updated references and paper

@@
 
-The remainder of this paper is structured as follows. @sec-data....
+The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate (done); 2) some detail about what the paper is about (done); 3) a clear gap that needs to be filled (done); 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper (done).
+A reader should be able to read only the introduction and know what was done, why, and what was found.
+Likely 3 or 4 paragraphs, or 10 per cent of total.
+
+As a response to the increase in gun violence in 2024 by 74% from 2023, the Toronto Police Services announced on June 27, 2024 at a press conference that they will increase their presence in areas impacted by high levels of gun violence [@citeincreasepolicetoronto].
+This concern was expressed because in a CBC article published in December 2023, they mentioned that there was a decrease in gun violence from 2019 to 2023 across Toronto [@citedropgunviolence].
+Deputy Chief Rob Johnson in the same conference also raised concerns on the increasing number of youth being perpetrators of gun violence [@citeyouthgunviolence].
+This raises the following questions: Which Toronto neighbourhoods are most impacted by shootings and homicides from 2018 to 2023?
+Where are police stations located relative to the shootings and homicides across Toronto?
+According to Statistics Canada, the most frequently used method to commit homicides is through shooting and so in our analysis we chose to analyze homicides alongside shootings [@citestatcan].
+
+In this paper, we analyzed data provided by Toronto Police Services and City Planning to investigate patterns where shootings and homicides take place and where police stations are located relative to where the crimes occur.
+Currently there is news coverage about increasing police visibility in areas with high gun violence but it is not known if the location of police stations in Toronto does impact the number of shootings and homicides that have taken place.
+
+In our findings,.....
+
+Our findings can also inform police divisions of neighbourhoods to be more present at based on how many cases has occurred in a neighbourhood.
+
+For the rest of the paper, the data section [@sec-data] will describe the datasets used and they were retrieved, our variables of interest used in our graphs and tables, and briefly about any variables that were constructed and the data cleaning process.
+In the results section [@sec-results], it will describe and show graphs and tables created based on our datasets and in the discussion section [@sec-discussion], it will provide an overview of what we did in our results, explain the meaning of our results and the implications it has on the real world, and provide potential areas of improvements of the paper and suggestions of future works.
 
@@
 For the cleaned dataset, we decided to filter for homicides and shootings and focus on the cases from 2018 to 2023, inclusive.
-The 2023 population projection in the dataset was provided by Environics Analytics where they counted only the residents in the neighbourhood [@citeneighbourhoodcrimerates].
+The 2023 population projection in the dataset was provided by Environics Analytics where they only included the number of residents in the neighbourhood [@citeneighbourhoodcrimerates].
 
-Each observation in the dataset represent one of Toronto's 158 neighbourhoods and its shootings and homicides from the years 2018 to 2023, inclusive that was recorded by the Toronto Police Services.
+Each observation in the dataset represent one of Toronto's 158 neighbourhoods and its shootings and homicides from the years 2018 to 2023, inclusive that was recorded by the Toronto Police Services as crimes occur in Toronto.
+The original dataset is updated annually by the Toronto Police Services in January of the following year.
 There are 1896 unique observations in total in the dataset.
@@
 ```
-## Number of Homicide and Shooting Cases in Toronto from 2018 to 2023 in Toronto
+## Number of Homicides and Shootings in Toronto from 2018 to 2023 in Toronto
 
@@
 # Discussion {#sec-discussion}
+In @sec-results, we focused on cases from 2018 to 2023 and looked at the number of homicides and shootings yearly and the six Toronto neighbourhoods that had the highest number of homicides and shootings using tables and graphs.
+Finally, we mapped where police facilities are located in Toronto alongside the number of homicides and shootings for each neighbourhood.
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Neighbourhoods with the highest number of shootings and homicides 
+- Glenfield-Jane Heights
+- Mount Olive-Silverstone-Jamestown	
+- Black Creek	
+- York University Heights	
+- Yorkdale-Glen Park
+- West Humber-Clairville
+
+https://www.npr.org/sections/money/2021/04/20/988769793/when-you-add-more-police-to-a-city-what-happens
+
+https://www.mcgilldaily.com/2023/02/more-policing-does-not-address-the-rise-of-violent-crime-in-canada/#:~:text=The%20NBER%20research%20found%20that,deeply%20harmful%20to%20marginalized%20communities
 
@@
 
-Weaknesses and next steps should also be included.
+We faced several limitations while conducting our analysis. We do not have information about when police facility were opened.
 
@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
 

### 2024-09-29 paper/paper.qmd
Finished paper

@@
 ---
-title: "Toronto Homicides and Shootings (To be Changed)"
+title: "High Number of Shootings and Homicides but Fewer Police Stations Around in Toronto Neighbourhoods"
 subtitle: "An analysis of police presence across Toronto neighbourhoods in relation to homicide and shooting cases (2018-2023)"
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Despite optimism in 2023 about gun violence continuing to decrease annually in Toronto, there was a sudden increase in gun violence at the beginning of 2024.
+This paper looks at patterns of where shootings and homicides occurred across all Toronto neighbourhoods alongside the location of police stations from 2018 to 2023.
+The results showed that the six neighbourhoods with the highest cases of shootings and homicides had below-average household income and smaller populations and there are fewer police stations around neighbourhoods with high numbers of shootings and homicides.
+These findings can inform where future police facilities can be built to ensure the safety of Toronto neighbourhoods, but further investigation is needed on how adding police facilities can impact crime cases in surrounding neighbourhoods over time."
 format:
@@
 
-The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate (done); 2) some detail about what the paper is about (done); 3) a clear gap that needs to be filled (done); 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper (done).
-A reader should be able to read only the introduction and know what was done, why, and what was found.
-Likely 3 or 4 paragraphs, or 10 per cent of total.
-
-As a response to the increase in gun violence in 2024 by 74% from 2023, the Toronto Police Services announced on June 27, 2024 at a press conference that they will increase their presence in areas impacted by high levels of gun violence [@citeincreasepolicetoronto].
-This concern was expressed because in a CBC article published in December 2023, they mentioned that there was a decrease in gun violence from 2019 to 2023 across Toronto [@citedropgunviolence].
-Deputy Chief Rob Johnson in the same conference also raised concerns on the increasing number of youth being perpetrators of gun violence [@citeyouthgunviolence].
+As a response to the increase in gun violence in 2024 by 74% from 2023, the Toronto Police Services announced on June 27 2024, at a press conference that they will increase their presence in areas impacted by high levels of gun violence [@citeincreasepolicetoronto].
+This concern was expressed because, in a CBC article published in December 2023, they mentioned that there was a decrease in gun violence from 2019 to 2023 across Toronto [@citedropgunviolence].
+Deputy Chief Rob Johnson in the same conference also raised concerns about the increasing number of youth being perpetrators of gun violence [@citeyouthgunviolence].
 This raises the following questions: Which Toronto neighbourhoods are most impacted by shootings and homicides from 2018 to 2023?
 Where are police stations located relative to the shootings and homicides across Toronto?
-According to Statistics Canada, the most frequently used method to commit homicides is through shooting and so in our analysis we chose to analyze homicides alongside shootings [@citestatcan].
+According to Statistics Canada, the most frequently used method to commit homicides is through shooting and in our analysis, we chose to analyze homicides alongside shootings [@citestatcan].
 
 In this paper, we analyzed data provided by Toronto Police Services and City Planning to investigate patterns where shootings and homicides take place and where police stations are located relative to where the crimes occur.
-Currently there is news coverage about increasing police visibility in areas with high gun violence but it is not known if the location of police stations in Toronto does impact the number of shootings and homicides that have taken place.
+Currently, there is news coverage about increasing police visibility in areas with high gun violence but it is not known if the location of police stations in Toronto does impact the number of shootings and homicides that have taken place.
+In our findings, our data showed that neighbourhoods with below-average household incomes and smaller populations have one of the highest numbers of shootings and homicides.
+Alongside this, we found that police stations were commonly not located in clusters of neighbourhoods with high numbers of shootings and homicides.
+These findings can inform police divisions of neighbourhoods to be more present at based on how many cases has occurred in a neighbourhood and figure out where to develop future police facilities.
 
-In our findings,.....
-
-Our findings can also inform police divisions of neighbourhoods to be more present at based on how many cases has occurred in a neighbourhood.
-
-For the rest of the paper, the data section [@sec-data] will describe the datasets used and they were retrieved, our variables of interest used in our graphs and tables, and briefly about any variables that were constructed and the data cleaning process.
-In the results section [@sec-results], it will describe and show graphs and tables created based on our datasets and in the discussion section [@sec-discussion], it will provide an overview of what we did in our results, explain the meaning of our results and the implications it has on the real world, and provide potential areas of improvements of the paper and suggestions of future works.
+For the rest of the paper, the data section [@sec-data] will describe the datasets used and how they were retrieved, our variables of interest used in our graphs and tables, and briefly about any variables that were constructed and the data cleaning process.
+The results section [@sec-results] will describe and show graphs and tables created based on our datasets and the discussion section [@sec-discussion] will provide an overview of what we did in our results, explain the meaning of our results and the implications it has on the real world, and provide potential areas of improvements of the paper and suggestions of future works.
 
@@
 @tbl-highestcasesshootingstable shows that Glenfield-Jane Heights has the highest number of shootings from 2018 to 2023 with 133 cases (4% of all shootings and homicides from 2018 to 2023) out of the 6 Toronto neighbourhoods with the highest number of shootings.
-Following this, Mount Olive-Silverstone-Jamestown has second highest number of shootings with 82 cases (3% of all shootings and homicides from 2018 to 2023).
+Following this, Mount Olive-Silverstone-Jamestown has the second highest number of shootings with 82 cases (3% of all shootings and homicides from 2018 to 2023).
 For the rest in descending order, Black Creek has 69 cases (2% of all shootings and homicides), York University Heights has 59 cases (2% of all shootings and homicides), Yorkdale-Glen Park has 58 cases (2% of all shootings and homicides), and Golfdale-Cedarbrae-Woburn has 54 cases (2% of all shootings and homicides).
@@
 However with other neighbourhoods with a high number of homicides such as West Humber-Clairville (1), Avondale (153), Glenfield-Jane Heights (25), York University Heights (27), and Eglinton East (138) has no police facilities in their neighbourhoods.
+@fig-homicides-map also shows that neighbourhoods in eastern, southern, and northwest Toronto have clusters with high numbers of homicides.
 
@@
 For the other three neighbourhoods, York University Heights	(27) and Golfdale-Cedarbrae-Woburn (141) have no police facilities located in the neighbourhood while Yorkdale-Glen Park (31) does.
+The neighbourhoods with the most number of shootings appears to be located in eastern and northern Toronto but there are fewer police stations located there compared to southern Toronto.
 
@@
 Finally, we mapped where police facilities are located in Toronto alongside the number of homicides and shootings for each neighbourhood.
+We saw that the yearly number of shootings was consistent with reports [@citedropgunviolence] indicating that gun violence was decreasing prior to 2024.
 
-## First discussion point {#sec-first-point}
+## Neighbourhoods with below average household incomes and smaller populations have the highest number of shootings and homicides
+@citewealthinequalities states that the average household income after-tax in 2020 for Toronto neighbourhoods is \$96000.
+Using our result [@tbl-highestcasestable] and further research from @citefindyourneighbourhood indicates that the six neighbourhoods with the highest number of shootings and homicides have below average household incomes after tax.
+Glenfield-Jane Heights, the neighbourhood with the highest number of cases, have an average after-tax household income in 2020 with \$76300 [@citefindyourneighbourhood].
+Mount Olive-Silverstone-Jamestown has an average household income of \$78200 while Black Creek has an average household income of \$68300 [@citefindyourneighbourhood].
+For the rest of the neighbourhoods, residents in York University Heights	make an average household income of \$77400 after tax while it is $81800 for Yorkdale-Glen Park residents and \$92300 for West Humber-Clairville [@citefindyourneighbourhood].
+From @tbl-highestcasestable, we also found that all neighbourhoods with high number of shootings and homicides have population sizes that made up less than 1% of Toronto's total population.
 
-Neighbourhoods with the highest number of shootings and homicides 
-- Glenfield-Jane Heights
-- Mount Olive-Silverstone-Jamestown	
-- Black Creek	
-- York University Heights	
-- Yorkdale-Glen Park
-- West Humber-Clairville
+## Police stations commonly not found in neighbourhoods with high number of shootings and homicides
 
-https://www.npr.org/sections/money/2021/04/20/988769793/when-you-add-more-police-to-a-city-what-happens
-
-https://www.mcgilldaily.com/2023/02/more-policing-does-not-address-the-rise-of-violent-crime-in-canada/#:~:text=The%20NBER%20research%20found%20that,deeply%20harmful%20to%20marginalized%20communities
-
-## Second discussion point
-
-## Third discussion point
+@fig-homicides-map and @fig-shootings-map showed that neighbourhoods with high number of shootings or homicides commonly did not have a police station located in them.
+However, there are some neighbourhoods with police stations located in the neighbourhood but still had a high number of cases.
+This indicates that having police stations located in the neighbourhood does not necessarily mean there is less crime cases in the specific neighbourhood.
+Although we found that clusters of neighbourhoods with higher number of cases had fewer police stations relative to the size of the cluster.
+In a study done by the National Bureau of Economic Research, they found that having more police presence in city can decrease the number of high-level crimes like homicides but this means more low-level crime arrests and traffic stops, which can hurt neighbourhoods with larger populations of marginalized individuals [@citemcgilldaily].
 
@@
 
-We faced several limitations while conducting our analysis. We do not have information about when police facility were opened.
+We faced several limitations while conducting our analysis.
+The datasets might not be completely accurate as human errors can occur when an entry is entered in the datasets.
+In the "Neighbourhood Crime Rates" dataset, the population is an approximation of the residents living in a neighbourhood at a certain point of time and does not consider commuters in the neighbourhoods.
+Also, the "Neighbourhood Crime Rates" dataset does not indicate if the shooting and/or homicide occurred along the borders of neighbourhoods and so the neighbourhood the case happened at is approximated.
+In the dataset we obtained on Toronto police facilities, it does not have information about which years each police facility were opened.
+This limited our analysis of police presence in relation to neighbourhood shootings and homicides.
 
 ## Next steps
+Results from our analysis can be used to inform decisions on where police divisions could increase their presence.
+Further investigations can look into how the appearance of police facilities impact the cases of shootings and/or homicides the neighbourhoods it is around over time.
+Another investigation can also look into the demographics of the neighbourhoods with high number of shootings and homicides.
 
